---
title: "Neural Networks, a better Human Nervous System?"
description: |
 This article gives an example of the uses of Neural Networks. 
output: 
  distill::distill_article
---

```{r load data and libraries, message=FALSE, echo=FALSE,warning=FALSE}
#https://archive.ics.uci.edu/ml/datasets/Las+Vegas+Strip
# r read csv from url
# allows you to directly download csv file from website
LV_data <- read.csv("https://archive.ics.uci.edu/ml/machine-learning-databases/00397/LasVegasTripAdvisorReviews-Dataset.csv", sep=';')

library(tidyverse)
library(caret)

draw_confusion_matrix <- function(cm) {
 
  layout(matrix(c(1,1,2)))
  par(mar=c(2,2,2,2))
  plot(c(100, 345), c(300, 450), type = "n", xlab="", ylab="", xaxt='n', yaxt='n')
  title('CONFUSION MATRIX', cex.main=2)
 
  # create the matrix
  rect(150, 430, 240, 370, col='#3F97D0')
  text(195, 435, 'Class1', cex=1.2)
  rect(250, 430, 340, 370, col='#F7AD50')
  text(295, 435, 'Class2', cex=1.2)
  text(125, 370, 'Predicted', cex=1.3, srt=90, font=2)
  text(245, 450, 'Actual', cex=1.3, font=2)
  rect(150, 305, 240, 365, col='#F7AD50')
  rect(250, 305, 340, 365, col='#3F97D0')
  text(140, 400, 'Class1', cex=1.2, srt=90)
  text(140, 335, 'Class2', cex=1.2, srt=90)
 
  # add in the cm results
  res <- as.numeric(cm$table)
  text(195, 400, res[1], cex=1.6, font=2, col='white')
  text(195, 335, res[2], cex=1.6, font=2, col='white')
  text(295, 400, res[3], cex=1.6, font=2, col='white')
  text(295, 335, res[4], cex=1.6, font=2, col='white')
 
  # add in the specifics
  plot(c(100, 0), c(100, 0), type = "n", xlab="", ylab="", main = "DETAILS", xaxt='n', yaxt='n')
  text(10, 85, names(cm$byClass[1]), cex=1.2, font=2)
  text(10, 70, round(as.numeric(cm$byClass[1]), 3), cex=1.2)
  text(30, 85, names(cm$byClass[2]), cex=1.2, font=2)
  text(30, 70, round(as.numeric(cm$byClass[2]), 3), cex=1.2)
  text(50, 85, names(cm$byClass[5]), cex=1.2, font=2)
  text(50, 70, round(as.numeric(cm$byClass[5]), 3), cex=1.2)
  text(70, 85, names(cm$byClass[6]), cex=1.2, font=2)
  text(70, 70, round(as.numeric(cm$byClass[6]), 3), cex=1.2)
  text(90, 85, names(cm$byClass[7]), cex=1.2, font=2)
  text(90, 70, round(as.numeric(cm$byClass[7]), 3), cex=1.2)
 
  # add in the accuracy information
  text(30, 35, names(cm$overall[1]), cex=1.5, font=2)
  text(30, 20, round(as.numeric(cm$overall[1]), 3), cex=1.4)
  text(70, 35, names(cm$overall[2]), cex=1.5, font=2)
  text(70, 20, round(as.numeric(cm$overall[2]), 3), cex=1.4)
} 

```

## Purpose

The purpose of this article is to display a use of Neural Networks as a machine learning model. Using the data from this [website](https://archive.ics.uci.edu/ml/machine-learning-databases/00397/LasVegasTripAdvisorReviews-Dataset.csv). Using Neural Networks, I will make predictions of whether the hotels in the observations have a tennis court on their facilities. 

## What's Neural Networks? 

The function of a neural network is to make predictions of what an observation will be. It takes an input data, and after many repetitions of mathematical algorithms, makes a prediction of what the output will be. It then compares those results with what actually occurs, then adjusts its process to become better. 

## The Model

In the code chunk below, I split the data in a training and testing set, then make predictions. 

```{r NNET, echo=TRUE, warning=FALSE, message=FALSE}
set.seed(1)

#To split the training / testing data
trainIndex <- createDataPartition(LV_data$Tennis.court, p = .6, list = FALSE, times = 1)

LV_dataTrain <- LV_data[ trainIndex,]
LV_dataTest  <- LV_data[-trainIndex,]

#Train the model

LV_dataNNET<- train(
  form = as.factor(Tennis.court) ~ Pool+Gym+Hotel.stars+Casino+Free.internet+Review.month+Score+Traveler.type+Nr..rooms,
  data = LV_dataTrain,
  trControl = trainControl(method = "cv", number = 10,
                           classProbs =  TRUE),
  method = "nnet",
  preProcess = c("center", "scale"),
  tuneLength = 5,
  trace=FALSE)

LV_dataNNET_Pred<-predict(LV_dataNNET,LV_dataTest,type="prob")

LV_dataNNETtestpred<-cbind(LV_dataNNET_Pred,LV_dataTest)

LV_dataNNETtestpred<-LV_dataNNETtestpred%>%
  mutate(prediction=if_else(NO>YES,"NO","YES"))
```

## Results

The table below displays what the model predicted vs. what actual occurred. Class2 indicates a Going Concern opinion issued, whereas Class1 is the opposite.

```{r confusion matrix,echo=FALSE}
NNETConfusion<-confusionMatrix(factor(LV_dataNNETtestpred$prediction),factor(LV_dataNNETtestpred$Tennis.court))

draw_confusion_matrix(NNETConfusion)
```

## How can Neural Networks be used in Accounting? 

Neural Networks can have many different uses. In the case of forensic accounting, it can be used to detect fraudulent transactions, which essentially enhances the efficiently and effectiveness of forensic investigations. 